---
title: "Using mascaRet"
author: "Fabrice Zaoui"
date: "April 08 2020"
output: html_document
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{A simple hydraulic example}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Short Description

**mascaRet** is an R package for the 1D shallow water and open-source code of the [Telemac](http://www.opentelemac.org) system.

Download the package from [GitHub](https://github.com/fzao/mascaRet) and then install and load it.

```{r inter}
library(mascaRet)
```

# Loading data

Change location to the example test and list the mascaRet files

```{r loading}
fpath <- system.file("extdata/Hydraulic", "mascaret.files", package="mascaRet")
mydir <- getwd()
datadir <- dirname(fpath)
setwd(datadir)
list.files()
```

- `mascaret.files` is the file listing all the data files
- `mascaret0.xcas` is the steering file of mascaRet (xml)
- `mascaret0_0.loi` and `mascaret0_1.loi` are graph files for boundary conditions
- `mascaret0.geo` is the file for the bathymetry information

# Create a mascaRet instantiation

This is the first thing to do before going further. An instance will be created and identified by a unique integer number

```{r create}
myMascProblem <- mascaRet_create()
print(myMascProblem)
```

The variable `myMascProblem` is important and will always be used hereafter

# Read hydraulic data from files

This is done with the help of `mascaret.files` and the import function

```{r read}
setwd(datadir)
verbose <- as.integer(0)  # silent mode (no listing)
error <- mascaRet_import(myMascProblem, verbose, 'mascaret.files')  # warning: myMascProblem must exist
print(error)
```

# Get

## Mesh info

Get the number of 1D nodes asking for the size of the abscissa vector

```{r getNode}
setwd(datadir)
varname <- "Model.B"  # bad abscissa name
index <- as.integer(1)  # 1D array
Size <- mascaRet_varsize(myMascProblem, varname, index)
print(Size)
print(mascaRet_error(myMascProblem))
varname <- "Model.X"  # abscissa name
index <- as.integer(1)  # 1D array
Size <- mascaRet_varsize(myMascProblem, varname, index)
nbnodes <- Size[1]  # number of nodes
print(nbnodes)
```

And get the values of bottom levels (m)

```{r getBottom}
setwd(datadir)
varname <- "Model.Zbot"  # variable name of the bottom values
Zbot <- vector("double", nbnodes)
for(i in 1:nbnodes) Zbot[i] <- mascaRet_get(myMascProblem, varname, i, 0, 0)
```

# Initialization

Initialize the hydraulic with default values for the water discharge (m3/s) and levels (m)

```{r init}
setwd(datadir)
q <- vector(mode = "numeric", length = nbnodes)  # disharges
z <- vector(mode = "numeric", length = nbnodes)  # water levels
error <- mascaRet_init(myMascProblem, q, z)
print(error)
```

# Computation

Define the time information for one step of calculation

```{r time}
tini <- 0.  # initial time (s)
tend <- 300.  # end of computation (s)
dt <- 300.  # time step (s)
```

Compute 1D shallow water for one step

```{r compute}
setwd(datadir)
error <- mascaRet_compute(myMascProblem, verbose, tini, tend, dt)
print(error)
```

# Graph results

Get the hydraulic state

```{r result}
Hydro <- mascaRet_hydro(myMascProblem)
print(names(Hydro))
```

Check if the hydraulic is a steady state

```{r check}
print(var(Hydro$Q))
```

And plot the water depth

```{r plot, echo=FALSE, message=FALSE}
#plot(Hydro$Z, xlab = "node number", ylab = "water levels (m)", type = "l")
library(plotly)
X <- seq(1, nbnodes)
Y <- Zbot
df <- data.frame(nodes=seq(1:nbnodes), bottom=Zbot, water=Hydro$Z)
fig <- plot_ly(df, x = ~nodes, y = ~water, type = 'scatter', mode = 'lines',
        line = list(color = 'rgba(0,128,255,1)'),
        showlegend = FALSE, name = 'Water levels')
fig <- fig %>% add_trace(y = ~bottom, type = 'scatter', mode = 'lines',
            fill = 'tonexty', fillcolor='rgba(77,166,255,0.2)', line = list(color = 'rgba(0,0,0,1)'),
            showlegend = FALSE, name = 'Bottom levels')
fig <- fig %>% layout(title = "Water Depth (m)",
         paper_bgcolor='rgb(255,255,255)', plot_bgcolor='rgb(229,229,229)',
         xaxis = list(title = "1D Nodes",
                      gridcolor = 'rgb(255,255,255)',
                      showgrid = TRUE,
                      showline = FALSE,
                      showticklabels = TRUE,
                      tickcolor = 'rgb(127,127,127)',
                      ticks = 'outside',
                      zeroline = FALSE),
         yaxis = list(title = "Levels (m)",
                      gridcolor = 'rgb(255,255,255)',
                      showgrid = TRUE,
                      showline = FALSE,
                      showticklabels = TRUE,
                      tickcolor = 'rgb(127,127,127)',
                      ticks = 'outside',
                      zeroline = FALSE))

fig
```

# Quit

Finally free the memory used by mascaRet and go back to the user directory

```{r return}
error <- mascaRet_delete(myMascProblem)
print(error)
setwd(mydir)
```