---
title: "Using the Water Quality module (Tracer) with mascaRet"
author: "Fabrice Zaoui"
date: "April 11 2020"
output: html_document
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Using the mascaRet R package with Tracer: a thermic example}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Short Description

**mascaRet** is an R package for the 1D shallow water and open-source code of the [Telemac](http://www.opentelemac.org) system. It uses a water quality module with different libraries (Tracer). This is a very simple case of temperature propagation in a flume. See the official documentation for details.

Download the package from [GitHub](https://github.com/fzao/mascaRet) and then install and load it.

```{r inter}
library(mascaRet)
```

# Loading data

Change location to the example test and list the mascaRet and Tracer files

```{r loading}
fpath <- system.file("extdata/Water_Quality", "mascaret.files", package="mascaRet")
mydir <- getwd()
datadir <- dirname(fpath)
setwd(datadir)
list.files()
```

- `mascaret.files` is the file listing all the data files
- `thermic.xcas` is the steering file of mascaRet (xml)
- `mascaret0_0.loi` and `mascaret0_1.loi` are graph files for boundary conditions
- `mascaret0.geo` is the file for the bathymetry information
- `mascaret0.lig` is the file for the initial hydraulic state
- `thermic.tracer_conc` is the file for the initial concentration of the constituents
- `thermic.tracer_loi` is boundary condition file for the constituent graph
- `thermic.tracer_parphy` is the the physical parameters file for thermic modelling 
- `thermic.tracer_meteo`is the file for the weather conditions

# Create a mascaRet instantiation

This is the first thing to do before going further. An instance will be created and identified by a unique integer number

```{r create}
myMascProblem <- mascaRet_create()
print(myMascProblem)
```

The variable `myMascProblem` is important and will always be used hereafter

# Read hydraulic data from files

This is done with the help of `mascaret.files` and the import function

```{r read}
setwd(datadir)
verbose <- as.integer(0)  # silent mode (no listing)
error <- mascaRet_import(myMascProblem, verbose, 'mascaret.files')  # warning: myMascProblem must exist
print(error)
```

# Initialization

## Hydraulic

Initialize the hydraulic with default values for the water discharge (m3/s) and levels (m) from file

```{r init}
setwd(datadir)
error <- mascaRet_initialization(myMascProblem, 'mascaret0.lig', verbose)
print(error)
```

## Tracer

Initialize the initial temperature with an array

First get the number of 1D nodes

Get the number of 1D nodes asking for the size of the abscissa vector

```{r get}
setwd(datadir)
varname <- "Model.X"  # abscissa name
index <- as.integer(1)  # 1D array
Size <- mascaRet_varsize(myMascProblem, varname, index)
nbnodes <- Size[1]  # number of nodes
print(nbnodes)
```

Then initialize the temperature with a default value of 10 degree Celsius

```{r temperature}
setwd(datadir)
ntracer <- 1
Tini <- matrix(10., nrow = nbnodes, ncol = ntracer)
error <- mascaRet_tracer_initialization(myMascProblem, verbose, Tini)
print(error)
```

# Computation

Define the time information for one step of calculation

```{r time}
tini <- 0.  # initial time (s)
tend <- 14400  # end of computation (s)
dt <- 3  # time step (s)
```

Compute 1D shallow water with Tracer

```{r compute}
setwd(datadir)
error <- mascaRet_compute(myMascProblem, verbose, tini, tend, dt)
print(error)
```

# Graph results

## Hydraulic

Get the hydraulic state

```{r result}
Hydro <- mascaRet_hydro(myMascProblem)
print(names(Hydro))
```

And plot the water levels

```{r plot}
plot(Hydro$Z, xlab = "node number", ylab = "water levels (m)", type = "l")
```

## Tracer

Get all the temperature values 

```{r resultTracer}
Temperature <- mascaRet_tracer_get(myMascProblem)
```

And plot

```{r plotTracer}
plot(Temperature[[1]], type = "l", xlab = "node number", ylab = "Temperature °C")
```

# Quit

Finally free the memory used by mascaRet and go back to the user directory

```{r return}
error <- mascaRet_delete(myMascProblem)
print(error)
setwd(mydir)
```